Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems

### 2015-08-17 Docs/Paper/LaTex/template.tex
review section 1

\textcolor{Red}{This article is organized by the following way: the section 2 gives a background of optimization, optimization with vZ, use of BMC technique with ESBMC, multi-cre ESBMC using OpenMP, ESBMC using optimized techniques of search to multiples cores with OpenMP. Section 3 describes the informal and formal mathematical modeling. Section 4 briefly describes th ILP and GA. The BMC method is based in SMT and is presented in section 5. The section 6 presents the paritioning model using vZ. The section 7 presents the experimental evaluation. The Section 8 discuss about the related works. The section 9 is presents the conclusions the future complementary works.}

\section{BACKGROUND}
\subsection{Optimization}
Optimization is the act of obtaining the best result (i.e., the optimal solution) under given circumstances [9]. In the design, construction, and maintenance of any engineering system, engineers have to make many technological and managerial decisions at several stages. The ultimate goal of all such decisions is either to minimize the effort required or to maximize the desired benefit. Because the effort required or the benefit desired in any practical situation can be expressed as a function of certain decision variables, optimization can be defined as the process of finding the conditions that give the maximum or minimum value of a function [9].

@@

\subsection{\textcolor{Red}{Optimization with Vz}}
A SMT solver is a tool that verifies formula's satisfiability based in a set of rules SMT. Z3 is a SMT solver made by Microsoft Research, freely available to the community, that supports a variety of theories. Z3 brings a solver DPLL together with a correspondence machine for quantifiers. Z3 is implemented in C++.
vZ is one of the tools implemented on the Z3 solver, it's base function is to optimize objectives within the logic context of constraints, it stands for the functions that make possible formulate optimized criteria.
vZ includes an incremental version of MaxRes to MaxSAT and a Simplex to solve numbers without defined patterns, in it's core MaxSAT is responsible to the restrictions while OptSMT optimize the linear arithmetic objectives.
Among all progress of Z3 using vZ the functions hightlights: maximize, minimize e assert-soft.

@@

\subsection{Bounded Model Checking with ESBMC}
Model checking refers to algorithms for exploring the state space of a transition system to determine if it obeys a speciﬁcation of its intended behavior [3], [4]. These algorithms can perform exhaustive exploration in a highly automatic way and, thus, have attracted much interest in industry. However, model-checking has been held back by the state explosion problem, in which the number of states in a system grows exponentially in the number of system components [10]. Much research has been devoted to mitigate this problem.

@@

\subsection{Multi-core ESBMC with OpenMP}
Nowadays, although the CPU used to perform tests usually has a modern multi-core architecture, with the ability to run several threads on different processing cores, ESBMC verification runs are still performed only in a single-core. For instance, if the processor has 8 processing cores available, only one is used for the verification and the others remain idle. There is a significant unused hardware resource during this process.

@@

\subsection{\textcolor{Red}{Multi-core ESBMC with OpenMP using Binary Search}}
\textcolor{Red}{The original parallelization was made by continuously forking the instances of the ESBMC until the first violation is find. Since OpenMP only returns from a parallelized loop when every forked thread finishes, some processor could remain idle for some time.}

@@

\section{Mathematical modeling}
The mathematical modeling was taken from [1], [2].
\subsection{Informal Model (or Assumptions)}
The informal model can be described by five characteristics. First, there is only one software context, i.e., there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning [17].

\subsection{Formal Model}
The inputs of the problem are: A directed simple graph $ G = (V,E) $, called the task graph of the system, is necessary. The vertices $ V = \{V_1,V_2,\dotso,V_n\} $ represent the nodes that are the components of the system that will be partitioned. The edges $ E $ represent communication between the components. Additionally, each node  $ V_i $ has a cost $ h(V_i) $ (or $ h_i $ of hardware (if implemented in hardware) and a cost $ s(s_i) $ (or $ s_i $ of software (if implemented in software). Finally, $ c(V_i,V_j) $ represents the communication cost between $ V_i $ and $ V_j $ if they are implemented in different contexts (hardware or software).

@@

\section{Partitioning problem using ILP-based,Genetic Algorithms}
\label{ILPGA}
The ILP and GA were taken from [5] and [6]. Both use slack variables in order to be possible to represent the constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to function ga of MATLAB [18]: the population size was set from 300 to 500, the Elite count changed from 2 (default value) to 50, and the number of Generations changed from 100* NumberOfVariables (default) to 75.

\section{Analysis of the partitioning problem using ESBMC}
ESBMC pseudocode shows the algorithm with the same restrictions and conditions placed on ILP and GA. Two values must be controlled to obtain the results and to perform the optimization. One is the initial software cost, as defined in Section III.G. The other is the halting condition (code violation) that stops the algorithm.

@@

\section{\textcolor{Red}{Analysis of the partitioning problem using vZ}}
The code below shows the conditions applied on ILP, GA and ESBMC converted to the vZ tool described in [34] and [35].

@@

\section{Experimental Evaluation}
ESBMC 1.24 running on a 64-bit Ubuntu 14.04.1 LTS operating system was used. Version 2.0.1 of Boolector SMT-solver [25] (freely available) was used as well. For the ILP and GA formulations, MATLAB R2013a from MathWorks with Parallel Computing Toolbox was used [18]. MATLAB is a dynamically typed high-level language known as the state-of-the-art mathematical software [19] and is widely used by the engineering community [20]. \textcolor{Red}{The built in Z3 tool vZ [34] available freely was also used. A parallel approach of the ESBMC sequential search ascending, descending and binary was implemented in c++11}. A desktop with 24GB of RAM and i7 (8-cores) from Intel with clock of 3.40 GHz was used. Each time was measured 3 times in GA (average taken) and just once in ESBMC and ILP. The reason is that GA times are not so close as ESBMC and ILP. A time out condition (TO) is reached when the running time is longer than 7,200 seconds. A memory out (MO) occurs when the tool reaches 24GB of memory. TABLE I. lists the benchmarks1.

@@

\section{Related Work}
Since the second half of the first decade of the 2000s, three main paths have been tracked to improve or to present alternative solutions to the optimization of HW-SW partitioning, i.e., to find the exact solution [2], to use heuristics to speed up performance time [1], and hybrid ones [22].

@@

In terms of SMT-based verification, most work is restricted to present the model, its modification to programming languages (e.g., C/C++ and Java), and the application to multi-thread algorithms or to embedded systems to check for program correctness. In [16] it presents a bounded model checker for C++ programs, which is an evolution of dealing with C programs and [14] uses the ESBMC model checker for embedded ANSI-C software. In [5] and [6] it was proven that it is possible to use ESBMC to solve HW-SW partitioning, but in a single core way. There are related studies focused on decreasing the verification time of model checkers by applying Swarm Verification [29], and modifications of internal search engines to support parallelism [30], but there is still the need for initiatives related to parallel SMT solvers [31]. Recently, the SMT solver Z3 has been extended to pose and solve optimization problems modulo theories [32].
\section{Conclusions}
Concerning the comparative tests, with the four techniques presented in this paper to solve HW-SW partitioning, it was evident that none of them is indicated to partition problems with more than 400 nodes. The computing time to solve the optimization problem reached some hours of execution on a standard desktop computer. If we consider less than 400 nodes, then it is possible to use ILP as the best solution provider. If the problem to be solved has 150 nodes or less, then ESBMC represents a feasible alternative. GA had an intermediate result in terms of performance, but the error presented from exact solution made it not acceptable to that kind of application. This error may be reduced by changing some parameters.

>>>>

\textcolor{blue}{Voces precisam citar as secoes usando o comando ref do latex. Para isso, cada secao deve conter um label.}
\textcolor{Red}{This article is organized as follows: Section~\ref{background} gives a background on optimization, optimization with vZ, use of BMC technique with ESBMC, multi-cre ESBMC using OpenMP, ESBMC using optimized techniques of search to multiples cores with OpenMP. Section 3 describes the informal and formal mathematical modeling. Section 4 briefly describes th ILP and GA. The BMC method is based on SMT and is presented in section 5. The section 6 presents the paritioning model using vZ. The section 7 presents the experimental evaluation. The Section 8 discusses about the related works. Section 9 presents the conclusions and the future work.}

%----------------------------------------------
\section{Background}
\label{background}
%----------------------------------------------

\textcolor{blue}{Introduzir uma breve descrição da seção.}

%----------------------------------------------
\subsection{Optimization}
%----------------------------------------------

Optimization is the act of obtaining the best result (i.e., the optimal solution) under given circumstances [9]. In the design, construction, and maintenance of any engineering system, engineers have to make many technological and managerial decisions at several stages. The ultimate goal of all such decisions is either to minimize the effort required or to maximize the desired benefit. Because the effort required or the benefit desired in any practical situation can be expressed as a function of certain decision variables, optimization can be defined as the process of finding the conditions that give the maximum or minimum value of a function [9].

@@

%----------------------------------------------
\subsection{\textcolor{Red}{Optimization with Vz}}
%----------------------------------------------

A SMT solver is a tool that verifies formula's satisfiability based in a set of rules SMT. Z3 is a SMT solver made by Microsoft Research, freely available to the community, that supports a variety of theories. Z3 brings a solver DPLL together with a correspondence machine for quantifiers. Z3 is implemented in C++.
vZ is one of the tools implemented on the Z3 solver, it's base function is to optimize objectives within the logic context of constraints, it stands for the functions that make possible formulate optimized criteria.
vZ includes an incremental version of MaxRes to MaxSAT and a Simplex to solve numbers without defined patterns, in it's core MaxSAT is responsible to the restrictions while OptSMT optimize the linear arithmetic objectives.
Among all progress of Z3 using vZ the functions hightlights: maximize, minimize e assert-soft.

@@

%----------------------------------------------
\subsection{Bounded Model Checking with ESBMC}
%----------------------------------------------

Model checking refers to algorithms for exploring the state space of a transition system to determine if it obeys a speciﬁcation of its intended behavior [3], [4]. These algorithms can perform exhaustive exploration in a highly automatic way and, thus, have attracted much interest in industry. However, model-checking has been held back by the state explosion problem, in which the number of states in a system grows exponentially in the number of system components [10]. Much research has been devoted to mitigate this problem.

@@

%----------------------------------------------
\subsection{Multi-core ESBMC with OpenMP}
%----------------------------------------------

Nowadays, although the CPU used to perform tests usually has a modern multi-core architecture, with the ability to run several threads on different processing cores, ESBMC verification runs are still performed only in a single-core. For instance, if the processor has 8 processing cores available, only one is used for the verification and the others remain idle. There is a significant unused hardware resource during this process.

@@

%----------------------------------------------
\subsection{\textcolor{Red}{Multi-core ESBMC with OpenMP using Binary Search}}
%----------------------------------------------

\textcolor{Red}{The original parallelization was made by continuously forking the instances of the ESBMC until the first violation is find. Since OpenMP only returns from a parallelized loop when every forked thread finishes, some processor could remain idle for some time.}

@@

%----------------------------------------------
\section{Mathematical modeling}
%----------------------------------------------

The mathematical modeling was taken from [1], [2].

%----------------------------------------------
\subsection{Informal Model (or Assumptions)}
%----------------------------------------------

The informal model can be described by five characteristics. First, there is only one software context, i.e., there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning [17].

%----------------------------------------------
\subsection{Formal Model}
%----------------------------------------------

The inputs of the problem are: A directed simple graph $ G = (V,E) $, called the task graph of the system, is necessary. The vertices $ V = \{V_1,V_2,\dotso,V_n\} $ represent the nodes that are the components of the system that will be partitioned. The edges $ E $ represent communication between the components. Additionally, each node  $ V_i $ has a cost $ h(V_i) $ (or $ h_i $ of hardware (if implemented in hardware) and a cost $ s(s_i) $ (or $ s_i $ of software (if implemented in software). Finally, $ c(V_i,V_j) $ represents the communication cost between $ V_i $ and $ V_j $ if they are implemented in different contexts (hardware or software).

@@

%----------------------------------------------
\section{Partitioning problem using ILP-based,Genetic Algorithms}
\label{ILPGA}
%----------------------------------------------

The ILP and GA were taken from [5] and [6]. Both use slack variables in order to be possible to represent the constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to function ga of MATLAB [18]: the population size was set from 300 to 500, the Elite count changed from 2 (default value) to 50, and the number of Generations changed from 100* NumberOfVariables (default) to 75.

%----------------------------------------------
\section{Analysis of the partitioning problem using ESBMC}
%----------------------------------------------

ESBMC pseudocode shows the algorithm with the same restrictions and conditions placed on ILP and GA. Two values must be controlled to obtain the results and to perform the optimization. One is the initial software cost, as defined in Section III.G. The other is the halting condition (code violation) that stops the algorithm.

@@

%----------------------------------------------
\section{\textcolor{Red}{Analysis of the partitioning problem using vZ}}
%----------------------------------------------

The code below shows the conditions applied on ILP, GA and ESBMC converted to the vZ tool described in [34] and [35].

@@

%----------------------------------------------
\section{Experimental Evaluation}
%----------------------------------------------

ESBMC 1.24 running on a 64-bit Ubuntu 14.04.1 LTS operating system was used. Version 2.0.1 of Boolector SMT-solver [25] (freely available) was used as well. For the ILP and GA formulations, MATLAB R2013a from MathWorks with Parallel Computing Toolbox was used [18]. MATLAB is a dynamically typed high-level language known as the state-of-the-art mathematical software [19] and is widely used by the engineering community [20]. \textcolor{Red}{The built in Z3 tool vZ [34] available freely was also used. A parallel approach of the ESBMC sequential search ascending, descending and binary was implemented in c++11}. A desktop with 24GB of RAM and i7 (8-cores) from Intel with clock of 3.40 GHz was used. Each time was measured 3 times in GA (average taken) and just once in ESBMC and ILP. The reason is that GA times are not so close as ESBMC and ILP. A time out condition (TO) is reached when the running time is longer than 7,200 seconds. A memory out (MO) occurs when the tool reaches 24GB of memory. TABLE I. lists the benchmarks1.

@@

%----------------------------------------------
\section{Related Work}
%----------------------------------------------

Since the second half of the first decade of the 2000s, three main paths have been tracked to improve or to present alternative solutions to the optimization of HW-SW partitioning, i.e., to find the exact solution [2], to use heuristics to speed up performance time [1], and hybrid ones [22].

@@

In terms of SMT-based verification, most work is restricted to present the model, its modification to programming languages (e.g., C/C++ and Java), and the application to multi-thread algorithms or to embedded systems to check for program correctness. In [16] it presents a bounded model checker for C++ programs, which is an evolution of dealing with C programs and [14] uses the ESBMC model checker for embedded ANSI-C software. In [5] and [6] it was proven that it is possible to use ESBMC to solve HW-SW partitioning, but in a single core way. There are related studies focused on decreasing the verification time of model checkers by applying Swarm Verification [29], and modifications of internal search engines to support parallelism [30], but there is still the need for initiatives related to parallel SMT solvers [31]. Recently, the SMT solver Z3 has been extended to pose and solve optimization problems modulo theories [32].

%----------------------------------------------
\section{Conclusions}
%----------------------------------------------

Concerning the comparative tests, with the four techniques presented in this paper to solve HW-SW partitioning, it was evident that none of them is indicated to partition problems with more than 400 nodes. The computing time to solve the optimization problem reached some hours of execution on a standard desktop computer. If we consider less than 400 nodes, then it is possible to use ILP as the best solution provider. If the problem to be solved has 150 nodes or less, then ESBMC represents a feasible alternative. GA had an intermediate result in terms of performance, but the error presented from exact solution made it not acceptable to that kind of application. This error may be reduced by changing some parameters.
